### 2024-11-06 Manuscript/main_ecography_R1.tex
Update main_ecography_R1.tex

The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity since 1500 is not greater than expected by chance (at least for certain metrics; Figure \ref{Fig:empirical_results}) because one would expect that extinctions since 1500 target specific areas of functional diversity space (i.e. birds with specific trait combinations).
This contrasts with studies that have shown that species with certain traits (e.g., larger species) are more likely to have gone extinct and thus that functional diversity loss is larger than through random extinctions (e.g. \citealt{sayol2021loss,Matthews2022}).
% TODO:
% This illustrates the importance of setting the question + metric
% For example, for the divergence metrics using probability density, hypervolume or dendrogram methods, we recovered changes in trait space that are smaller than expected by chance for the species present before 1500 and larger than expected by change for the current species.
% This could lead to the interpretation that the trait space was not changed under any specific mechanism before 1500 (i.e. null mechanism) but that since then, specific mechanisms have been reducing specific areas of the trait space (e.g. filtering mechanism).
% However, when measuring divergence using the Rao's quadratic entropy based on distance matrices, we did not find that.
This could partly be due to the proportion of passerine extinctions in Hawaii.
Although most global bird species are (or were, if accounting for anthropogenic extinctions) passerines (approximatively 58\%), the majority of global known anthropogenic extinctions (approx. 75\%) have been non-passerines, which tend to be larger and possess more unique morphological traits (thus resulting in relatively large reductions in functional diversity).
However, in Hawaii, where the same proportion of known species were passerines (approx. 58\%), the opposite is true: the majority of extinctions (approx. 60\%) have been of passerine species.
It is important to note that we are not implying that there existed little morphological variation within extinct Hawaiian passerines - indeed, many of these extinctions involved the honeycreepers, a famous adaptive radiation involving substantial evolution of beak morphology \citep{Walther2022} - but simply that this morphological variation is less than that observed across all birds.
But these results, given that the proportion of known Hawaiian species that are / were passerines roughly matches the proportion of extinct species that were passerines,  also indicate that these extinct species were more functionally similar than expected, thus relatively \textit{increasing} trait diversity (i.e. relative to random extinctions).
In addition, it is worth stressing that these analyses included marine species (which constitute less than 4\% of  known extinctions in Hawaii, while making up 11\% of the prehistoric assemblage) and different results may have been found if focusing exclusively on terrestrial birds.

>>>>

The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity since 1500 is not greater than expected by chance (at least for certain metrics; Figure \ref{Fig:empirical_results}) because one would expect that extinctions since 1500 target specific areas of functional diversity space (i.e. birds with specific trait combinations).
This contrasts with studies that have shown that species with certain traits (e.g., larger species) are more likely to have gone extinct and thus that functional diversity loss is larger than through random extinctions (e.g. \citealt{sayol2021loss,Matthews2022}).
This illustrates the importance of positing the mechanism and process of interest as well as the pattern to capture it before analysing the results.
For example, for the divergence metrics using probability density, hypervolume or dendrogram methods, we recovered changes in trait space that are smaller than expected by chance for the species present before 1500 and larger than expected by change for the current species.
This could lead to the interpretation that the trait space was not changed under any specific mechanism before 1500 (i.e. null mechanism) but that since then, specific mechanisms have been reducing specific areas of the trait space (e.g. filtering mechanism).
However, when using a different metric (say the Rao's quadratic entropy based on distance matrices), the results would indicate that a reduction of specific areas of the trait space could have also occured before 1500 leading to very different interpretation of the results.

Our results also indicate that although most global bird species are (or were, if accounting for anthropogenic extinctions) passerines (approximatively 58\%), the majority of global known anthropogenic extinctions (approx. 75\%) have been non-passerines, which tend to be larger and possess more unique morphological traits (thus resulting in relatively large reductions in functional diversity).
However, in Hawaii, where the same proportion of known species were passerines (approx. 58\%), the opposite is true: the majority of extinctions (approx. 60\%) have been of passerine species.
It is important to note that we are not implying that there existed little morphological variation within extinct Hawaiian passerines - indeed, many of these extinctions involved the honeycreepers, a famous adaptive radiation involving substantial evolution of beak morphology \citep{Walther2022} - but simply that this morphological variation is less than that observed across all birds.
But these results, given that the proportion of known Hawaiian species that are / were passerines roughly matches the proportion of extinct species that were passerines,  also indicate that these extinct species were more functionally similar than expected, thus relatively \textit{increasing} trait diversity (i.e. relative to random extinctions).
In addition, it is worth stressing that these analyses included marine species (which constitute less than 4\% of  known extinctions in Hawaii, while making up 11\% of the prehistoric assemblage) and different results may have been found if focusing exclusively on terrestrial birds.
